---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading Packages and Libraries
``` {r libraries}
install.packages("ggplot2")
library(ggplot2)
```

## Loading and preprocessing the data
``` {r LoadingAndPreprocessing, echo = TRUE}
activityData <- read.csv("data\\activity.csv")
activityData$date <- as.Date(activityData$date, format = "%Y-%m-%d")

```

## What is mean total number of steps taken per day?

``` {r meanSteps, echo = TRUE}
#For this part of the assignment ignore the missing values in the dataset.
activityDataWithoutMissingValues <- na.omit(activityData)
aggregateStepsByDate <- aggregate(steps ~ date, activityDataWithoutMissingValues, sum)

#Make a histogram of the total number of steps taken each day
qplot(aggregateStepsByDate$steps, geom="histogram", xlab = "Total Number of Steps taken each day", main = "Histogram - Total number of steps taken each day (with 300 binwidth)", binwidth = 300)

```
``` {r summary, echo = TRUE}
summary(aggregateStepsByDate)

```
You can see from the above table that the mean is `r mean(aggregateStepsByDate$steps)`, while the median is `r median(aggregateStepsByDate$steps)`.


## What is the average daily activity pattern?

``` {r averageDailyPattern, echo = TRUE}

aggregateAvgStepsByDate <- aggregate(steps ~ interval, activityDataWithoutMissingValues, mean)
plot(aggregateAvgStepsByDate$interval, aggregateAvgStepsByDate$steps, type ="l", main = "Time Series of the 5 min interval and avg. # of steps accross all days")

aggregateAvgStepsByDate[which.max(aggregateAvgStepsByDate$steps), ]

```
On Average accross all the days in the dataset, the 5 minute interval  `r aggregateAvgStepsByDate[which.max(aggregateAvgStepsByDate$steps),"interval"]` contains the maximun number of steps.

## Imputing missing values

The total number of rows with NAs:

```{r ImputtingMissingVal, echo = TRUE}

sum(is.na(activityData))

```

```{r FillInMissingVal, echo = TRUE}
#Using the Avg. # of steps for the missing 5 minute interval, fill in the missing value

activityDataWithCompleteValues <- merge(activityData, aggregateAvgStepsByDate, by = c("interval"))

activityDataWithCompleteValues$steps <- ifelse(is.na(activityDataWithCompleteValues$steps.x) == TRUE,activityDataWithCompleteValues$steps.y, activityDataWithCompleteValues$steps.x)

activityDataWithCompleteValues <- subset(activityDataWithCompleteValues, select=-c(steps.x,steps.y))

# Make a histogram
aggregateStepsByDate_WithFill <- aggregate(steps ~ date, activityDataWithCompleteValues, sum)

#Make a histogram of the total number of steps taken each day
qplot(aggregateStepsByDate_WithFill$steps, geom="histogram", xlab = "Total Number of Steps taken each day", main = "Histogram - Total number of steps taken each day (with 300 binwidth) With Fill", binwidth = 300)

```

``` {r FillInMissingValSummary, echo = TRUE}
summary(aggregateStepsByDate_WithFill)

```

When compared to the original histogram prior to filling in the NA values, the avg stayed the same and the median only changed slightly. No significant changes to the histogram are found.

## Are there differences in activity patterns between weekdays and weekends?

```{r WeekendComparison, echo = TRUE}

activityDataWithCompleteValues$DayType <- factor(ifelse(weekdays(activityDataWithCompleteValues$date, abbr = TRUE) %in% c("Sat", "Sun"), "weekend", "weekday"))

aggregateAvgStepsByDateAndDateType <- aggregate(steps ~ interval+DayType, activityDataWithCompleteValues, mean)

qplot(interval,
      steps,
      data = aggregateAvgStepsByDateAndDateType,
      facets = .~ DayType,
      geom = "line")

```

